---
title: "Creating landscapes"
author: "Gabriel Nakamura"
date: "2025-04-06"
output: html_document
---


This file is used to create landscapes in which simulation processess in 
    gen3sis will be running
    
# Packages and functions

```{r}
require(gen3sis)
require(raster)
require(scales) # Package to rescale variables, vectors and ranges.

```

# Geographical landscape and coordinates

Here we define the geographical limits, the size of our landscape and the 
    temporal variability 
    
```{r}
# Fixing geographical limits.
latmin <- -10
latmax <- 0
longmin <- -5
longmax <- 5

# Defining the size of our landscape.
Ncol <- 10
Nrow <- 10
Nsites <- Ncol*Nrow
# N sites Ncol*Nrow = 100.

# Defining the temporal variability of the landscape.
Ntimesteps <- 1000

# Mapping site-center coordinates: lat , long.
coordinates <- data.frame(lat = rep(seq(latmax, latmin, length.out=Nrow), each=Ncol), long = rep(seq(longmin, longmax, length.out=Ncol), times=Nrow))
# It's needed ordering latitude from maximum to minimum for that the matrix/raster be properly built.

# Checking coordinates.
coordinates$lat
coordinates$long

```

# Environmental characteristics of the landscape

Creating the environmental features of the landscape 

## Plain ladscape 

```{r}
# Creating arrays of environmental variables. Each array has three dimensions.
temp_array <- array(NA, dim=c(Nrow, Ncol, Ntimesteps))

# Creating environmental data.frames.
temp_df <- matrix(NA, nrow=(Nrow*Ncol), ncol=(length(c(Ncol,Nrow))+Ntimesteps))

# Creating a vector for the names to be used to name input files (one per time step)
timesteps_names <- vector(mode="character", length=Ntimesteps)
```

## Building the landscape 

```{r}
# Here the landscape is built.

# Making the landscape.
# Temporal loop.
for(timestep in 1:Ntimesteps){
  # Counter for elements.
  counting <- 1

  # Naming the timesteps.
  timesteps_names[timestep] <- paste("X", timestep, "", sep="")

  # Loops for the spatial dimensions.
  for(i in 1:Nrow){
    for(j in 1:Ncol){
      # Temperature is defined as a function of the latitude and with a normal random component.
      # The mean temperature scale is linear and fixed as 30ºC in Equator and as -30ºC in poles.
      # The standard deviation is defined as 2ºC.
      #cat(coordinates$lat[counting], coordinates$long[counting], "\n")

      temp_array[i, j, timestep] <- 30 - 0.667*abs(coordinates$lat[counting]) + rnorm(1, sd=2)


      # Saving the environmental variables in a dataframe format for distance matrices.
      temp_df[counting, 2+timestep] <- temp_array[i, j, timestep]
      counting <- counting + 1
    }
  }
}
# Associating coordinates to raster.
temp_df[, 1] <- coordinates$long
temp_df[, 2] <- coordinates$lat
```

# Saving landscape 

```{r}
# Saving landscape.

# Define the work directory.
# simulation_path <- "~/Dropbox/ada_simulations/plain_landscape"
simulation_path <- here::here("gen3sis_simulations", "landscape_plain")
dir.create(simulation_path)
setwd(simulation_path)

# Create a list of rasters with landscape information.
landscapes_list <- list()
for (timestep in 1:Ntimesteps){
  # Create a raster from the landscape associated to each time step.
  temp_raster <- rasterFromXYZ(temp_df[, c(1, 2, timestep+2)])
  # prec_raster <- rasterFromXYZ(prec_dataframe[, c(1,2, timestep+2)])

  # Add the raster of each time step to the list.
  landscapes_list$temp <- c(landscapes_list$temp, temp_raster)
  # landscapes_list$prec <- c(landscapes_list$prec, prec_raster)
}

# Save the landscapes list as a rds file.
saveRDS(landscapes_list, file.path(here::here("gen3sis_simulations", "landscape_plain", "landscape_list.rds")))
```

Saving the objects in your hardrive

```{r}
################################################################################
# Importing the raster list.
landscapes <- readRDS(file.path(simulation_path, "landscape_list.rds"))
landscapes$temp[[1]]
# Using the rds format the data is imported as a list, then it's not needed modify it.

################################################################################
# Defining a function cost.

# NULL FUNCTION COST: all the sites are equivalents.
# source: the index of source site (origin).
# habitable_src: habitability of source site.
# dest: the index of destiny site.
# habitable_dest: habitability of destiny site.
cost_function_null <- function(source, habitable_src, dest, habitable_dest){
  # Return 1 for any pair of sites.
  return(1)
}

################################################################################
# Creating the distances to input landscape in the simulation.
create_input_landscape(landscapes = landscapes, 
                       overwrite_output = TRUE, 
                       # The cost function to calculate the distances in the landscape.
                       cost_function = cost_function_null,
                       # Directory name to save the files in.
                       output_directory = file.path(simulation_path, "landscapes"),
                       # All surrounding sites from a focal site: 4, 8 or 16 neighboring sites.
                       directions = 8,
                       calculate_full_distance_matrices = T
                       # Calculate full distance matrices or only surrounding sites distances.
                       # In the second case the distance between other sites is calculated
                       # at each time step of simulation. Then:
                       # TRUE: large storage and higher speed computing.
                       # FALSE: lower storage and lower speed computing.
)

################################################################################
# Finally, it's really important documenting the landscape you created. It is
# doing writing the METADATA.txt file within the landscape folder. It's
# automatically created by the function `create_input_landscape`.

```



