---
title: "Extracting co-occurrences from simulations"
author: "Gabriel Nakamura"
date: "2025-04-04"
output: html_document
---

# Context

# Functions and packages

```{r}
source(here::here("R_functions", "phylo_gen3sis_tools_1.R"))
source(here::here("R_functions", "get_node_site_matrix_GN.R"))

simulation_main <- here::here("gen3sis_simulations", "gen3sis_output_1", "config_presence")
simulation_output <- here::here("gen3sis_simulations", "gen3sis_output_1", "output_communities", "output_sim1.rds")

```


# Scenarios

## Scenario 1

Extracting community composition at all timesteps, phylogenetic tree and a dataframe 
   with the labels of species and their birth time

```{r}

# main function to extract all the information - this takes a few minutes (5min in my windows 24GB  pc)
res_sim_1 <- 
  get_node_site_matrix(simulation_path = simulation_main, save = TRUE, output_file = simulation_output)

```

Extracting community composition at time zero and a matrix with community composition 
    of nodes (occurrence of nodes in different sites ) - RUN FROM HERE

```{r}

res_sim_comm <- readRDS(simulation_output)

# node and species composition matrix
occ_nodes_spp <- res_sim_comm$node_site_matrix
occ_nodes_spp <- res_sim_1$node_site_matrix

# phylogeny
phylo_sim <- res_sim_comm$tree
phylo_sim <- res_sim_1$tree
# data frame with species and node information
df_nodes_spp <- res_sim_comm$nameIndexGuide
df_nodes_spp <- res_sim_1$nameIndexGuide

# coordinates of the sites 
df_coords <- res_sim_comm$coordinates
df_coords <- res_sim_1$coordinates



# only the presence of species in time zero in each site
occ_time_zero <- occ_nodes_spp[, match(phylo_sim$tip.label, colnames(occ_nodes_spp))]

# only the presence of nodes in each site
occ_nodes_only <- occ_nodes_spp[, -match(phylo_sim$tip.label, colnames(occ_nodes_spp))]

hist(colSums(occ_time_zero)) # just to check out the distribution of occurrences of species 


```


# Scenario 3 - less coexistence


```{r}
source(here::here("R_functions", "phylo_gen3sis_tools_1.R"))
source(here::here("R_functions", "get_node_site_matrix_GN.R"))

simulation_main <- here::here("gen3sis_simulations", "gen3sis_output_presence_less_coex", "config_presence_less_coex")
simulation_output <- here::here("gen3sis_simulations", "gen3sis_output_1", "output_communities", "output_sim1.rds")

```

Extracting communities

```{r}
# main function to extract all the information - this takes a few minutes (5min in my windows 24GB  pc)
res_sim_2 <- 
  get_node_site_matrix(simulation_path = simulation_main, save = FALSE)

occ_nodes_spp <- res_sim_2$node_site_matrix

# phylogeny
phylo_sim <- res_sim_2$tree

# data frame with species and node information
df_nodes_spp <- res_sim_2$nameIndexGuide

# coordinates of the sites 
df_coords <- res_sim_2$coordinates



# only the presence of species in time zero in each site
occ_time_zero <- occ_nodes_spp[, match(phylo_sim$tip.label, colnames(occ_nodes_spp))]

# only the presence of nodes in each site
occ_nodes_only <- occ_nodes_spp[, -match(phylo_sim$tip.label, colnames(occ_nodes_spp))]

hist(colSums(occ_time_zero)) # just to check out the distribution of occurrences of species 

```



