---
title: "Multiple simulation scenario difusion"
author: "Gabriel Nakamura"
date: "2025-04-12"
output: html_document
---

# Scenario 1 - High species richness and 100 grids

This scenario is based on the config file `config_presence.R`. The simulation 
    uses a landscape with 100 grids and the parameters are set in a way to produce
    high levels of species richness.

```{r}
library(gen3sis)

# informing the local of configuration file used in this simulation
path_config <- here::here("gen3sis_simulations", "config", "config_presence", "config_presence.R")
config_obj <- create_input_config(config_file = path_config)
path_output <- here::here("gen3sis_simulations_grid100_high_species")
# informing the landscape used in this simulation scenario
landscape_scenario1 <- file.path(here::here("gen3sis_simulations", "landscape_plain", "landscapes"))

# number of simulations
reps_sim <- seq(1:100) 

# running the simulations
res_sim_disp_only <- 
  lapply(1:length(reps_sim), function(x){
    config_obj$gen3sis$general$random_seed <- reps_sim[x]
    gen3sis::run_simulation(config = config_obj, 
                            landscape = landscape_scenario1, 
                            output_directory = here::here(path_output, paste("gen3sis_output_dispersion", x, sep = "_")))
  }
  )


```

Extracting the simulation results from folders and saving the community composition 
    and phylogenetic tree in an output folder

```{r}
# extracting the results 

source(here::here("R_functions", "phylo_gen3sis_tools_1.R"))
source(here::here("R_functions", "get_node_site_matrix_GN.R"))

simulation_main <- here::here("gen3sis_simulations_grid100_high_species")

res_all_sim_disp <- 
  lapply(1:length(reps_sim), function(x){
    
    dir_temp <- 
      dir(here::here(simulation_main,
                     paste("gen3sis_output_dispersion_", x, sep = ""),
                     "default_config")
      )
    
    simulation_path_res <- 
      here::here(simulation_main,
                     paste("gen3sis_output_dispersion_", x, sep = ""),
                     "default_config", dir_temp)
      
    
    simulation_output <- 
      here::here(simulation_main,
                 "gen3sis_output_dispersion_res_100_highdiv",
                 paste("output_dispersion", x, ".rds", sep = "")
                 )
    
    res_sim <- 
      get_node_site_matrix(simulation_path = simulation_path_res, 
                           save = TRUE, 
                           output_file = simulation_output
      )
  })


```


# Scenario 2 - Low species richness and 100 grids

We reduced the number of species by reducing the number of maximum coexistences
    allowed in the configuration file
    

```{r}
# informing the local of configuration file used in this simulation
path_config <- here::here("gen3sis_simulations", "config", "config_presence_less_coex", "config_presence_less_coex.R")
config_obj <- create_input_config(config_file = path_config)
path_output <- here::here("gen3sis_simulations_low_diversity_grid100")
# informing the landscape used in this simulation scenario\
landscape_scenario1 <- file.path(here::here("gen3sis_simulations", "landscape_plain", "landscapes"))

# number of simulations
reps_sim <- seq(1:100) 

# running the simulations
res_sim_disp_only <- 
  lapply(1:length(reps_sim), function(x){
    config_obj$gen3sis$general$random_seed <- reps_sim[x]
    gen3sis::run_simulation(config = config_obj, 
                            landscape = landscape_scenario1, 
                            output_directory = here::here(path_output, paste("gen3sis_output_dispersion_lowDiv_grid100", x, sep = "_")))
  }
  )

```

Extracting communities

```{r}
# extracting the results 

source(here::here("R_functions", "phylo_gen3sis_tools_1.R"))
source(here::here("R_functions", "get_node_site_matrix_GN.R"))

simulation_main <- here::here("gen3sis_simulations_low_diversity_grid100")

res_all_sim_disp <- 
  lapply(1:length(reps_sim), function(x){
    
    dir_temp <- 
      dir(here::here(simulation_main,
                     paste("gen3sis_output_dispersion_lowDiv_grid100", x, sep = "_"),
                     "default_config")
      )
    
    simulation_path_res <- 
      here::here(simulation_main,
                     paste("gen3sis_output_dispersion_lowDiv_grid100", x, sep = "_"),
                     "default_config", dir_temp)
      
    
    simulation_output <- 
      here::here(simulation_main,
                 "gen3sis_output_lowdiv_100",
                 paste("output_dispersion_low_spp_100_", x, ".rds", sep = "")
                 )
    
    res_sim <- 
      get_node_site_matrix(simulation_path = simulation_path_res, 
                           save = TRUE, 
                           output_file = simulation_output
      )
  })

```

# Scenario 3 - High species richness and 900 grids

```{r}
library(gen3sis)

# informing the local of configuration file used in this simulation
path_config <- here::here("gen3sis_simulations", "config", "config_presence", "config_presence.R")
config_obj <- create_input_config(config_file = path_config)
path_output <- here::here("gen3sis_simulations_high_diversity_grid900")
# informing the landscape used in this simulation scenario
landscape_scenario1 <- file.path(here::here("gen3sis_simulations", "landscape_plain_900", "landscapes"))

# number of simulations
reps_sim <- seq(1:100) 

# running the simulations
res_sim_disp_only <- 
  lapply(1:length(reps_sim), function(x){
    config_obj$gen3sis$general$random_seed <- reps_sim[x]
    gen3sis::run_simulation(config = config_obj, 
                            landscape = landscape_scenario1, 
                            output_directory = here::here(path_output, paste("gen3sis_output_dispersion", x, sep = "_")))
  }
  )


```

Extracting communities

```{r}
# extracting the results 

source(here::here("R_functions", "phylo_gen3sis_tools_1.R"))
source(here::here("R_functions", "get_node_site_matrix_GN.R"))

simulation_main <- here::here("gen3sis_simulations_high_diversity_grid900")

res_all_sim_disp <- 
  lapply(1:length(reps_sim), function(x){
    
    dir_temp <- 
      dir(here::here(simulation_main,
                     paste("gen3sis_output_dispersion", x, sep = "_"),
                     "default_config")
      )
    
    simulation_path_res <- 
      here::here(simulation_main,
                     paste("gen3sis_output_dispersion", x, sep = "_"),
                     "default_config", dir_temp)
      
    
    simulation_output <- 
      here::here(simulation_main,
                 "gen3sis_output_high_diversity_grid900",
                 paste("output_dispersion_low_spp_100_", x, ".rds", sep = "")
                 )
    
    res_sim <- 
      get_node_site_matrix(simulation_path = simulation_path_res, 
                           save = TRUE, 
                           output_file = simulation_output
      )
  })

```
# Scenario 4 - Low species richness and 900 grids

Here we reduce the number of species by reducing the alpha diversity in simulations

```{r}
library(gen3sis)

# informing the local of configuration file used in this simulation
path_config <- here::here("gen3sis_simulations", "config", "config_presence_less_coex", "config_presence_less_coex.R")
config_obj <- create_input_config(config_file = path_config)
path_output <- here::here("gen3sis_simulations_low_diversity_grid900")
# informing the landscape used in this simulation scenario
landscape_scenario1 <- file.path(here::here("gen3sis_simulations", "landscape_plain_900", "landscapes"))

# number of simulations
reps_sim <- seq(1:100) 

# running the simulations
res_sim_disp_only <- 
  lapply(1:length(reps_sim), function(x){
    config_obj$gen3sis$general$random_seed <- reps_sim[x]
    gen3sis::run_simulation(config = config_obj, 
                            landscape = landscape_scenario1, 
                            output_directory = here::here(path_output, paste("gen3sis_output_dispersion", x, sep = "_")))
  }
  )


```

Extracting communities

```{r}
# extracting the results 

source(here::here("R_functions", "phylo_gen3sis_tools_1.R"))
source(here::here("R_functions", "get_node_site_matrix_GN.R"))

simulation_main <- here::here("gen3sis_simulations_low_diversity_grid900")

res_all_sim_disp <- 
  lapply(1:length(reps_sim), function(x){
    
    dir_temp <- 
      dir(here::here(simulation_main,
                     paste("gen3sis_output_dispersion", x, sep = "_"),
                     "default_config")
      )
    
    simulation_path_res <- 
      here::here(simulation_main,
                     paste("gen3sis_output_dispersion", x, sep = "_"),
                     "default_config", dir_temp)
      
    
    simulation_output <- 
      here::here(simulation_main,
                 "gen3sis_output_low_diversity_grid900",
                 paste("output_dispersion_low_spp_100_", x, ".rds", sep = "")
                 )
    
    res_sim <- 
      get_node_site_matrix(simulation_path = simulation_path_res, 
                           save = TRUE, 
                           output_file = simulation_output
      )
  })

```

