---
title: "Summary statistics of simulated communities"
author: "Gabriel Nakamura"
date: "2025-06-11"
output: html_document
---

# Initial setup

Reading libraries

```{r}
library(ggplot2)
library(patchwork)
library(ggridges)
library(dplyr)
```

Reading output files that consist of .rds files with lists that are the output of the 
    simulations. 
    
- $node_site_matrix: A matrix object with grids in the rows and nodes in the columns
    with the information of presence (1) or absence (0) of the nodes in all the 
    grids

- $tree: a phylo object containing the phylogenetic tree that resulted from simulations

- $nameIndexGuide: A data frame object containing three columns.
  - id: the name of the species in the original tree
  - height: the time step in which the species originated by cladogenetic process
  - timestep: the timestep of the simulation process in which the species originated 
  

Summary statistics for simulated communities with high number of species and 
    100 grids

Reading files from simulations

```{r}

rds_files_highdiv_100 <- list.files(path = here::here("gen3sis_simulations_grid100_high_species", "gen3sis_output_dispersion_res_100_highdiv"), pattern = "\\.rds$", full.names = TRUE)
rds_files_highdiv_900 <- list.files(path = here::here("gen3sis_simulations_high_diversity_grid900", "gen3sis_output_high_diversity_grid900"), pattern = "\\.rds$", full.names = TRUE)
rds_files_lowdiv_100 <- list.files(path = here::here("gen3sis_simulations_low_diversity_grid100", "gen3sis_output_lowdiv_100"), pattern = "\\.rds$", full.names = TRUE)
rds_files_lowdiv_900 <- list.files(path = here::here("gen3sis_simulations_low_diversity_grid900", "gen3sis_output_low_diversity_grid900"), pattern = "\\.rds$", full.names = TRUE)

res_highdiv_100 <- lapply(rds_files_highdiv_100, function(x) readRDS(x))
res_highdiv_900 <- lapply(rds_files_highdiv_900, function(x) readRDS(x))
res_lowdiv_100 <- lapply(rds_files_lowdiv_100, function(x) readRDS(x))
res_lowdiv_900 <- lapply(rds_files_lowdiv_900, function(x) readRDS(x))

```

# Number of species 

Computing the number of species for all simulations

```{r}


richness_highdiv_100 <- 
  unlist(lapply(res_highdiv_100, function(x){
    length(x$tree$tip.label)
  }))


richness_highdiv_900 <- 
  unlist(lapply(res_highdiv_900, function(x){
    length(x$tree$tip.label)
  }))

richness_lowdiv_100 <- 
  unlist(lapply(res_lowdiv_100, function(x){
    length(x$tree$tip.label)
  }))

richness_lowdiv_900 <- 
  unlist(lapply(res_lowdiv_900, function(x){
    length(x$tree$tip.label)
  }))


df_rich_simulations <-
  data.frame(richness = c(richness_highdiv_100, richness_highdiv_900, richness_lowdiv_100, richness_lowdiv_900),
             scenario = rep(c("High Div/100", "High Div/900", "Low Div/100", "Low Div/900"), each = length(richness_highdiv_100)))

```

# Plotting richness in each scenario

```{r}
# Plot
plot_richness_all_scenario <- 
  ggplot(df_rich_simulations, aes(x = richness, y = scenario, fill = scenario)) +
  geom_density_ridges(
    scale = 1, 
    alpha = 0.4, 
    rel_min_height = 0.01, 
    color = NA
  ) +
  geom_jitter(
    aes(color = scenario),
    height = 0.2, 
    size = 1.5, 
    alpha = 0.6, 
    show.legend = FALSE
  ) +
  scale_fill_manual(values = c("High Div/100" = "#E69F00", 
                               "High Div/900" = "#56B4E9",
                               "Low Div/100" = "#009E73", 
                               "Low Div/900" = "#D55E00")
                    ) +
scale_color_manual(values = c("High Div/100" = "#E69F00",
                              "High Div/900" = "#56B4E9",
                              "Low Div/100" = "#009E73",
                              "Low Div/900" = "#D55E00")
                   ) +
  theme_minimal() +
  labs(title = "", x = "Richness", y = "") +
  theme(
    legend.position = "none",
    axis.line = element_line(color = "black"),          # Add black axis lines
    axis.title.x = element_text(color = "black", size = 12),       # Make x-axis title black
    axis.title.y = element_text(color = "black"),        # Make y-axis title black
    axis.text.x = element_text(color = "black", size = 11),
    axis.text.y = element_text(color = "black", size = 12),
    axis.ticks.y = element_line(color = "black"), 
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

# saving the plot
#ggsave(here::here("output", "FigS1_richness.png"),
#       width = 9, height = 7,
#       dpi = 600, plot = plot_richness_all_scenario)

ggsave(here::here("output", "FigS1_richness.pdf"), 
       plot = plot_richness_all_scenario, 
       width = 9,
       height = 7)
```



# Abundance distribution 

```{r}

freq_list_highdiv_100 <- 
  lapply(res_highdiv_100, function(x){
    colSums(x$node_site_matrix) 
  })  

df_freq_list_highdiv_100 <- 
  lapply(freq_list_highdiv_100, function(x){
    data.frame(freq = x)
  })

freq_list_highdiv_900 <- 
  lapply(res_highdiv_900, function(x){
    colSums(x$node_site_matrix) 
  })  

df_freq_list_highdiv_900 <- 
  lapply(freq_list_highdiv_900, function(x){
    data.frame(freq = x)
  })

freq_list_lowdiv_100 <- 
  lapply(res_lowdiv_100, function(x){
    colSums(x$node_site_matrix) 
  })  

df_freq_list_lowdiv_100 <- 
  lapply(freq_list_lowdiv_100, function(x){
    data.frame(freq = x)
  })

freq_list_lowdiv_900 <- 
  lapply(res_lowdiv_900, function(x){
    colSums(x$node_site_matrix) 
  })  

df_freq_list_lowdiv_900 <- 
  lapply(freq_list_lowdiv_900, function(x){
    data.frame(freq = x)
  })




```

Joining all simulations categories in a single data frame

```{r}
vec_reps_spp_high_100 <- unlist(lapply(df_freq_list_highdiv_100, function(x) nrow(x)))
simul_category <- rep(paste("simul", 1:100), times = vec_reps_spp_high_100)
df_occ_high_100 <- data.frame(Abundance = do.call(rbind, df_freq_list_highdiv_100), group = simul_category) 

vec_reps_spp_high_900 <- unlist(lapply(df_freq_list_highdiv_900, function(x) nrow(x)))
simul_category <- rep(paste("simul", 1:100), times = vec_reps_spp_high_900)
df_occ_high_900 <- data.frame(Abundance = do.call(rbind, df_freq_list_highdiv_900), group = simul_category) 

vec_reps_spp_low_100 <- unlist(lapply(df_freq_list_lowdiv_100, function(x) nrow(x)))
simul_category <- rep(paste("simul", 1:100), times = vec_reps_spp_low_100)
df_occ_low_100 <- data.frame(Abundance = do.call(rbind, df_freq_list_lowdiv_100), group = simul_category) 

vec_reps_spp_low_900 <- unlist(lapply(df_freq_list_lowdiv_900, function(x) nrow(x)))
simul_category <- rep(paste("simul", 1:100), times = vec_reps_spp_low_900)
df_occ_low_900 <- data.frame(Abundance = do.call(rbind, df_freq_list_lowdiv_900), group = simul_category) 

```


# Plotting abundance frequency


```{r}

# high diversity with 100 grids
plot_dens_high_100 <- 
  ggplot(df_occ_high_100, aes(x = freq, group = group)) +
  geom_density(color = "gray50", size = 0.6, alpha = 0.4) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line = element_line(color = "black", size = 0.4),
    axis.ticks = element_line(color = "black", size = 0.3)
  ) + 
  labs(x = "", y = "Density")


# high diversity with 900 grids
plot_dens_high_900 <- 
  ggplot(df_occ_high_900, aes(x = freq, group = group)) +
  geom_density(color = "gray50", size = 0.6, alpha = 0.4) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line = element_line(color = "black", size = 0.4),
    axis.ticks = element_line(color = "black", size = 0.3)
  ) + 
  labs(x = "", y = "")

# low diversity with 100 grids
plot_dens_low_100 <- 
  ggplot(df_occ_low_100, aes(x = freq, group = group)) +
  geom_density(color = "gray50", size = 0.6, alpha = 0.4) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line = element_line(color = "black", size = 0.4),
    axis.ticks = element_line(color = "black", size = 0.3)
  ) + 
  labs(x = "Number of grids occupied", y = "Density")

# low diversity with 900 grids
plot_dens_low_900 <- 
  ggplot(df_occ_low_900, aes(x = freq, group = group)) +
  geom_density(color = "gray50", size = 0.6, alpha = 0.4) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line = element_line(color = "black", size = 0.4),
    axis.ticks = element_line(color = "black", size = 0.3)
  ) + 
  labs(x = "Number of grids occupied", y = "")



# joining and saving graphics

all_density <- 
  plot_dens_high_100+plot_dens_high_900+plot_dens_low_100+plot_dens_low_900+
  patchwork::plot_annotation(tag_levels = "a")+
  patchwork::plot_layout(nrow = 2)&
  theme(
    plot.margin = margin(5.5,8,5.5,5.5,"pt")
  )

#ggsave(here::here("output", "FigSX_all_density_occupancy.png"),
#       width = 9, height = 7,
#       dpi = 600, plot = all_density)

ggsave(here::here("output", "FigSX_all_density_occupancy.pdf"),
       width = 9, height = 7,
       plot = all_density)


```
